EldersAIR initial Micropem data work

### 2020-09-03 micropem_data_cleaning.Rmd
data cleaning/checking updates

# Format combined file from above; combine with other PM data below
npt_micropem <- initial_data %>% 
  separate(time, c("trash1", "time_pm"), sep = " ", remove = FALSE) %>% 
  separate(list_files, c("mp_file_name", "trash2"), sep = ".xlsx", remove = FALSE) %>%
  mutate(date_pm = ymd(date),
         pm_personal = as.numeric(pm),
         area = "NPT",
         mp_file_name = as.character(mp_file_name),
         file_type = "xlsx") %>% 
  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(mp_file_name, area, datetime_pm, date_pm, time_pm, pm_personal, file_type,
         temp:shutdown_reason) %>% 
  filter(!is.na(pm_personal)) %>% 
  ungroup() %>% 
  arrange(area, mp_file_name, datetime_pm)

@@

npt_micropem_new <- npt_micropem %>%   
  group_by(area, mp_file_name) %>% 
  # new variables summarizing pm2.5 within each home
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         first_datetime = first(datetime_pm),
         last_datetime = last(datetime_pm),
         first_datetime = ymd_hms(first_datetime),
         last_datetime = ymd_hms(last_datetime),
         diff_datetime = interval(first_datetime, last_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm_personal, na.rm = TRUE),
         pm_sd = sd(pm_personal, na.rm = TRUE),
         pm_min = min(pm_personal, na.rm = TRUE),
         pm_median = median(pm_personal, na.rm = TRUE),
         pm_max = max(pm_personal, na.rm = TRUE),
         # breaking sampling period into days
         obs = row_number(),
         sampling_day = if_else(obs < 1441, 1, 0),
         sampling_day = if_else(obs < 2881 & obs > 1440, 2, sampling_day),
         sampling_day = if_else(obs < 4321 & obs > 2880, 3, sampling_day)) %>%
  # calculate pm means for each sampling day
  group_by(area, mp_file_name, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm_personal, na.rm = TRUE),
         pm_sampling_day = sampling_day,
         day_of_week = weekdays(date_pm)) %>%
  mutate(start_datetime_pm = ymd_hms(start_datetime_pm),
         pm_start_diff = interval(first_datetime, start_datetime_pm),
         pm_start_diff_hrs = pm_start_diff/3600) %>% 
  ungroup() %>% 
  select(-obs, -sampling_day) %>% 
  arrange(area, home, sampling_visit)

@@

pm_desc <<- npt_micropem %>% 
  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
  arrange(desc(pm_personal)) %>% 
  select(datetime_pm, pm_personal, pm_sampling_day)
pm_time_plot

@@

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "51", location = "NPT", visit = 2)
head(gaps, 10)
head(pm_desc, 10)
tail(pm_desc, 10)
# filter data for a single home
check_data <- elders_micropem %>% 
  filter(area == "NPT" & home_winter_id == "37" & sampling_visit == 1) %>% 
  filter(pm_personal < 0)
```

>>>>

# Format combined file from above; combine with other PM data below
npt_micropem <- initial_data %>% 
  separate(time, c("trash1", "time_pm"), sep = " ", remove = FALSE) %>% 
  separate(list_files, c("mp_file_name", "trash2"), sep = ".xlsx", remove = FALSE) %>%
  mutate(date_pm = ymd(date),
         pm_personal = as.numeric(pm),
         area = "NPT",
         mp_file_name = as.character(mp_file_name),
         file_type = "xlsx") %>% 
  unite("datetime", c("date_pm", "time_pm"), sep = " ", remove = FALSE) %>% 
  mutate(datetime_pm = ymd_hms(datetime)) %>% 
  select(mp_file_name, area, datetime_pm, date_pm, time_pm, pm_personal, file_type,
         temp:shutdown_reason) %>% 
  filter(!is.na(pm_personal)) %>% 
  filter(!is.na(rh)) %>% 
  ungroup() %>% 
  arrange(area, mp_file_name, datetime_pm)

@@

npt_micropem_new <- npt_micropem %>%   
  group_by(area, mp_file_name) %>% 
  # new variables summarizing pm2.5 within each home
  mutate(pm_total_observations = n(),
         # new datetime values to calculate sampling interval
         first_datetime = first(datetime_pm),
         last_datetime = last(datetime_pm),
         first_datetime = ymd_hms(first_datetime),
         last_datetime = ymd_hms(last_datetime),
         diff_datetime = interval(first_datetime, last_datetime),
         # total sampling interval, based on first and last datetimes
         pm_sample_interval = diff_datetime/86400,
         # expected number of observations with the sampling interval
         expected_obs = pm_sample_interval*24*60,
         # difference between actual and expected observations
         actual_vs_expected_diff = pm_total_observations - expected_obs,
         # indicator variable if actual vs expected obs > 10 or greater
         actual_vs_expected = if_else(actual_vs_expected_diff < 10 &
                                      actual_vs_expected_diff > -10, 1, 0),
         pm_mean = mean(pm_personal, na.rm = TRUE),
         pm_sd = sd(pm_personal, na.rm = TRUE),
         pm_min = min(pm_personal, na.rm = TRUE),
         pm_median = median(pm_personal, na.rm = TRUE),
         pm_max = max(pm_personal, na.rm = TRUE),
         # breaking sampling period into days
         sampling_day = if_else(datetime_pm - first_datetime < 86400, 1, 0),
         sampling_day = if_else(datetime_pm - first_datetime >= 86400 & datetime_pm - first_datetime < 172800, 2, sampling_day),
         sampling_day = if_else(datetime_pm - first_datetime > 172800, 3, sampling_day)) %>%
  filter(sampling_day != 0) %>% 
  # calculate pm means for each sampling day
  group_by(area, mp_file_name, sampling_day) %>% 
  mutate(pm_mean_daily = mean(pm_personal, na.rm = TRUE),
         pm_sampling_day = sampling_day,
         day_of_week = weekdays(date_pm)) %>%
  group_by(area, mp_file_name) %>% 
  mutate(start_datetime_pm = ymd_hms(start_datetime_pm),
         pm_start_diff = interval(first_datetime, start_datetime_pm),
         pm_start_diff_hrs = pm_start_diff/3600) %>% 
  ungroup() %>% 
  select(-sampling_day) %>% 
  arrange(area, home_winter_id, sampling_visit)

@@

pm_desc <<- npt_micropem %>% 
  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
  arrange(desc(pm_personal)) %>% 
  select(datetime_pm, pm_personal, pm_sampling_day)

pm_negs <<- npt_micropem %>% 
  filter(home_winter_id == id & area == location & sampling_visit == visit) %>% 
  filter(pm_personal < 0) %>% 
  arrange(desc(pm_personal)) %>% 
  select(datetime_pm, pm_personal, pm_sampling_day)

pm_time_plot

@@

# Change function input to look at specific home
# Run all of these lines together
pm_time_plot_function(id = "65", location = "NPT", visit = 1)
head(gaps, 10)
head(pm_desc, 10)
tail(pm_desc, 10)

# filter data for a single home
check_data <- npt_micropem %>% 
  filter(area == "NPT" & home_winter_id == "48" & sampling_visit == 2) %>% 
  filter(pm_personal < 0)
```
